---
title: "Comprehensive Data Visualization in R: Univariate & Multivariate Analysis"
author: "Data Analytics Team"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
    number_sections: true
    theme: flatly
    highlight: tango
    code_folding: show
    df_print: paged
    fig_width: 8
    fig_height: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
# Global knitr options for clean, consistent rendering
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 8,
  fig.height = 5,
  dpi = 100
)

# Clean professional color palette used throughout the document
main_color   <- "#2C3E50"   # dark slate (primary)
accent_color <- "#2980B9"   # blue (secondary)
highlight    <- "#E67E22"   # orange (highlight/contrast)
palette_cat  <- c("#2C3E50", "#2980B9", "#27AE60", "#E67E22", "#C0392B", "#8E44AD")
```

```{r install-packages, include=FALSE}
# Install any missing packages required for this document (run once).
# This chunk is silent (include=FALSE) and only installs what's missing.
required_packages <- c(
  "ggplot2", "dplyr", "tidyr", "moments", "GGally", "corrplot",
  "reshape2", "vcd", "scatterplot3d", "plotly", "MASS",
  "pheatmap", "RColorBrewer", "knitr", "kableExtra"
)

installed <- rownames(installed.packages())
to_install <- setdiff(required_packages, installed)
if (length(to_install) > 0) {
  install.packages(to_install, repos = "https://cloud.r-project.org")
}
```

```{r load-libraries, message=FALSE, warning=FALSE}
# Load all libraries needed for the analyses below
library(ggplot2)       # Grammar-of-graphics plotting
library(dplyr)         # Data manipulation
library(tidyr)         # Data reshaping
library(moments)       # Skewness / kurtosis
library(GGally)        # Scatterplot matrices (ggpairs)
library(corrplot)      # Correlation visualizations
library(reshape2)      # melt() for heatmaps
library(vcd)           # Mosaic plots
library(scatterplot3d) # Static 3D scatter plots
library(plotly)        # Interactive plots
library(MASS)          # parcoord() for parallel coordinates
library(pheatmap)      # Clustered heatmaps
library(RColorBrewer)  # Color palettes
library(knitr)         # Tables
library(kableExtra)    # Enhanced tables

# Consistent ggplot2 theme applied to all plots in this document
theme_set(
  theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", size = 14, color = "#2C3E50"),
      panel.grid.minor = element_blank(),
      legend.position = "right"
    )
)
```

# Introduction

This document demonstrates a comprehensive set of **univariate**,
**bivariate**, and **multivariate** visualization techniques in R, using
both **Base R** and **ggplot2** (plus a few specialized packages). The
built-in `mtcars` dataset is used throughout for consistency and
reproducibility.

```{r data-overview}
# Load and inspect the mtcars dataset
data(mtcars)

# Create a copy with categorical variables properly labeled as factors
mtcars_df <- mtcars %>%
  mutate(
    cyl_f  = factor(cyl, labels = c("4 cyl", "6 cyl", "8 cyl")),
    am_f   = factor(am, labels = c("Automatic", "Manual")),
    vs_f   = factor(vs, labels = c("V-shaped", "Straight")),
    gear_f = factor(gear),
    carb_f = factor(carb)
  )

str(mtcars_df)
kable(head(mtcars_df, 6), caption = "First 6 rows of the mtcars dataset")
```

------------------------------------------------------------------------

# Univariate Analysis: Numerical Variables

We focus on `mpg` (miles per gallon) and `hp` (horsepower) as our
primary continuous variables of interest.

## Summary Statistics

```{r univariate-summary-stats}
# Compute a rich set of descriptive statistics for a numeric vector
summary_stats <- function(x, name = "variable") {
  data.frame(
    Variable = name,
    N        = length(x),
    Mean     = round(mean(x), 3),
    Median   = round(median(x), 3),
    SD       = round(sd(x), 3),
    Variance = round(var(x), 3),
    Min      = round(min(x), 3),
    Max      = round(max(x), 3),
    Range    = round(max(x) - min(x), 3),
    IQR      = round(IQR(x), 3),
    Skewness = round(moments::skewness(x), 3),
    Kurtosis = round(moments::kurtosis(x), 3)
  )
}

stats_table <- rbind(
  summary_stats(mtcars_df$mpg, "mpg"),
  summary_stats(mtcars_df$hp,  "hp"),
  summary_stats(mtcars_df$wt,  "wt")
)

kable(stats_table, caption = "Descriptive statistics for key numeric variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

**Interpretation notes:** Skewness near 0 indicates symmetry (positive =
right-tailed, negative = left-tailed). Excess kurtosis near 0 (moments
package returns raw kurtosis, so \~3 = normal-like tails) indicates
normal-like peakedness; values above 3 indicate heavier tails
(leptokurtic), below 3 indicate lighter tails (platykurtic).

## Histograms with Density Overlays

### Base R

```{r hist-base-r}
# Base R histogram of mpg with a density curve overlaid
hist(mtcars_df$mpg,
     breaks = 10,
     freq = FALSE,                     # freq=FALSE lets us overlay a density
     col = "#2980B9",
     border = "white",
     main = "Distribution of MPG (Base R)",
     xlab = "Miles per Gallon",
     ylab = "Density")
lines(density(mtcars_df$mpg), col = "#E67E22", lwd = 3)  # density overlay
rug(mtcars_df$mpg, col = "#2C3E50")                        # rug of raw values
legend("topright", legend = c("Density"), col = "#E67E22", lwd = 3, bty = "n")
```

### ggplot2

```{r hist-ggplot2}
# ggplot2 histogram + density overlay (density scaled to match histogram counts)
ggplot(mtcars_df, aes(x = mpg)) +
  geom_histogram(aes(y = after_stat(density)), bins = 10,
                  fill = accent_color, color = "white", alpha = 0.85) +
  geom_density(color = highlight, linewidth = 1.1) +
  labs(title = "Distribution of MPG (ggplot2)",
       x = "Miles per Gallon", y = "Density") +
  theme_minimal()
```

## Boxplots with Jittered Points

### Base R

```{r boxplot-base-r}
# Base R boxplot with individual jittered points overlaid
boxplot(mtcars_df$mpg,
        main = "Boxplot of MPG (Base R)",
        ylab = "Miles per Gallon",
        col = "#2980B9", border = "#2C3E50")
set.seed(42)
points(jitter(rep(1, nrow(mtcars_df)), amount = 0.05), mtcars_df$mpg,
       pch = 19, col = adjustcolor("#E67E22", alpha.f = 0.7))
```

### ggplot2

```{r boxplot-ggplot2}
# ggplot2 boxplot with jittered raw data points on top
ggplot(mtcars_df, aes(x = "", y = mpg)) +
  geom_boxplot(fill = accent_color, alpha = 0.4, width = 0.4, outlier.shape = NA) +
  geom_jitter(width = 0.08, color = highlight, alpha = 0.8, size = 2) +
  labs(title = "Boxplot of MPG with Jittered Points (ggplot2)",
       x = NULL, y = "Miles per Gallon") +
  theme_minimal()
```

## Density Plots with Rug

```{r density-rug-ggplot2}
# ggplot2 density plot with a rug showing individual observations
ggplot(mtcars_df, aes(x = mpg)) +
  geom_density(fill = accent_color, alpha = 0.5, color = main_color, linewidth = 1) +
  geom_rug(color = highlight, alpha = 0.7, linewidth = 0.8) +
  labs(title = "Density Plot of MPG with Rug", x = "Miles per Gallon", y = "Density") +
  theme_minimal()
```

## Q-Q Plots with Normality Tests

### Base R

```{r qqplot-base-r}
# Base R Q-Q plot to visually assess normality of mpg
qqnorm(mtcars_df$mpg, main = "Normal Q-Q Plot of MPG (Base R)",
       col = "#2980B9", pch = 19)
qqline(mtcars_df$mpg, col = "#E67E22", lwd = 2)  # reference line
```

### ggplot2

```{r qqplot-ggplot2}
# ggplot2 Q-Q plot using stat_qq / stat_qq_line
ggplot(mtcars_df, aes(sample = mpg)) +
  stat_qq(color = accent_color, size = 2) +
  stat_qq_line(color = highlight, linewidth = 1) +
  labs(title = "Normal Q-Q Plot of MPG (ggplot2)",
       x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
```

```{r shapiro-test}
# Shapiro-Wilk test for normality (H0: data are normally distributed)
shapiro_mpg <- shapiro.test(mtcars_df$mpg)
shapiro_mpg

# Interpretation helper
cat(sprintf("\nShapiro-Wilk p-value = %.4f -> %s\n",
            shapiro_mpg$p.value,
            ifelse(shapiro_mpg$p.value > 0.05,
                   "Fail to reject normality (consistent with a normal distribution)",
                   "Reject normality (data deviate significantly from normal)")))
```

------------------------------------------------------------------------

# Univariate Analysis: Categorical Variables

We use `cyl` (number of cylinders) and `am` (transmission type) as our
categorical variables of interest.

## Frequency Tables and Proportions

```{r freq-tables}
# Frequency counts and proportions for number of cylinders
cyl_freq <- mtcars_df %>%
  count(cyl_f) %>%
  mutate(Proportion = round(n / sum(n), 3),
         Percentage = paste0(round(n / sum(n) * 100, 1), "%"))

kable(cyl_freq, caption = "Frequency table: Number of Cylinders") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)
```

## Bar Plots with Value Labels

### Base R

```{r barplot-base-r}
# Base R bar plot with counts labeled above each bar
counts <- table(mtcars_df$cyl_f)
bp <- barplot(counts, col = palette_cat[1:3], border = "white",
              main = "Count of Cars by Cylinder (Base R)",
              ylab = "Count", ylim = c(0, max(counts) * 1.2))
text(bp, counts, labels = counts, pos = 3, font = 2)  # value labels above bars
```

### ggplot2

```{r barplot-ggplot2}
# ggplot2 bar plot with count labels on top of each bar
ggplot(mtcars_df, aes(x = cyl_f, fill = cyl_f)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count)),
            vjust = -0.5, fontface = "bold") +
  scale_fill_manual(values = palette_cat[1:3]) +
  labs(title = "Count of Cars by Cylinder (ggplot2)",
       x = "Number of Cylinders", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Pie Charts

> **Caution:** Pie charts are generally discouraged in modern data
> visualization practice because humans are less accurate at comparing
> angles/areas than bar lengths. A bar plot (above) is almost always a
> better choice. The pie chart below is included only for completeness /
> comparison purposes.

```{r piechart-base-r}
# Base R pie chart (use sparingly - see caution note above)
pie(counts, labels = paste0(names(counts), "\n(", counts, ")"),
    col = palette_cat[1:3],
    main = "Cylinder Distribution (Pie Chart - use with caution)")
```

------------------------------------------------------------------------

# Bivariate Analysis

## Scatter Plots with Regression Line and Correlation Annotation

```{r scatter-regression}
# Compute correlation for annotation
cor_test <- cor.test(mtcars_df$wt, mtcars_df$mpg)
cor_label <- sprintf("r = %.2f, p = %.4f", cor_test$estimate, cor_test$p.value)

# ggplot2 scatter plot with linear regression line and correlation annotation
ggplot(mtcars_df, aes(x = wt, y = mpg)) +
  geom_point(color = accent_color, size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", color = highlight, fill = "grey80", se = TRUE) +
  annotate("text", x = max(mtcars_df$wt) * 0.75, y = max(mtcars_df$mpg) * 0.95,
           label = cor_label, size = 4.5, fontface = "italic", color = main_color) +
  labs(title = "MPG vs. Weight with Linear Fit",
       x = "Weight (1000 lbs)", y = "Miles per Gallon") +
  theme_minimal()
```

## Boxplots by Group with Violin Overlays

```{r boxplot-violin-group}
# Combined violin + boxplot to show distribution shape AND summary stats by group
ggplot(mtcars_df, aes(x = cyl_f, y = mpg, fill = cyl_f)) +
  geom_violin(alpha = 0.4, trim = FALSE) +
  geom_boxplot(width = 0.15, fill = "white", alpha = 0.8, outlier.color = highlight) +
  scale_fill_manual(values = palette_cat[1:3]) +
  labs(title = "MPG Distribution by Cylinder Count (Violin + Boxplot)",
       x = "Number of Cylinders", y = "Miles per Gallon") +
  theme_minimal() +
  theme(legend.position = "none")
```

## Grouped Bar Plots

```{r grouped-barplot}
# Grouped bar plot: average mpg by cylinder count and transmission type
summary_group <- mtcars_df %>%
  group_by(cyl_f, am_f) %>%
  summarise(mean_mpg = mean(mpg), .groups = "drop")

ggplot(summary_group, aes(x = cyl_f, y = mean_mpg, fill = am_f)) +
  geom_col(position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(aes(label = round(mean_mpg, 1)),
            position = position_dodge(width = 0.8), vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = c(main_color, highlight)) +
  labs(title = "Average MPG by Cylinder Count and Transmission Type",
       x = "Number of Cylinders", y = "Average MPG", fill = "Transmission") +
  theme_minimal()
```

## Mosaic Plots and Contingency Tables

```{r contingency-table}
# Contingency table: cylinders x transmission type
cont_table <- table(mtcars_df$cyl_f, mtcars_df$am_f)
kable(as.data.frame.matrix(cont_table),
      caption = "Contingency Table: Cylinders x Transmission Type") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Chi-square test of independence
chisq_result <- suppressWarnings(chisq.test(cont_table))
chisq_result
```

```{r mosaic-plot}
# Mosaic plot visualizing the same contingency table
mosaic(cont_table,
       main = "Mosaic Plot: Cylinders x Transmission Type",
       shade = TRUE, legend = TRUE)
```

------------------------------------------------------------------------

# Multivariate Analysis

## Scatterplot Matrix with GGally

```{r ggpairs-matrix, fig.width=10, fig.height=8}
# Scatterplot matrix showing pairwise relationships, distributions, and
# correlations across several numeric + categorical variables at once
GGally::ggpairs(
  mtcars_df,
  columns = c("mpg", "hp", "wt", "qsec"),
  aes(color = cyl_f, alpha = 0.7),
  title = "Scatterplot Matrix: MPG, HP, Weight, and Quarter-Mile Time"
) +
  scale_color_manual(values = palette_cat[1:3]) +
  scale_fill_manual(values = palette_cat[1:3]) +
  theme_minimal(base_size = 9)
```

## Correlation Matrix and Correlation Tests

```{r correlation-matrix}
# Correlation matrix across key numeric variables
num_vars <- mtcars_df %>% select(mpg, disp, hp, drat, wt, qsec)
cor_matrix <- cor(num_vars, method = "pearson")

kable(round(cor_matrix, 2), caption = "Pearson Correlation Matrix") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE)

# Correlation significance test for a specific pair as an example
cor.test(mtcars_df$hp, mtcars_df$mpg, method = "pearson")
```

## Correlation Heatmap with Values Displayed

```{r correlation-heatmap}
# corrplot heatmap with correlation coefficients printed inside each cell
corrplot(cor_matrix,
         method = "color",
         type = "upper",
         addCoef.col = "black",       # print correlation values
         number.cex = 0.8,
         tl.col = main_color, tl.srt = 45,
         col = colorRampPalette(c("#C0392B", "white", "#2980B9"))(200),
         title = "Correlation Heatmap (mtcars numeric variables)",
         mar = c(0, 0, 2, 0))
```

## Color-Coded Scatter Plots (3+ Dimensions)

```{r multidim-scatter}
# Scatter plot encoding 4 dimensions at once:
# x = wt, y = mpg, color = cyl (categorical), size = hp (continuous)
ggplot(mtcars_df, aes(x = wt, y = mpg, color = cyl_f, size = hp)) +
  geom_point(alpha = 0.75) +
  scale_color_manual(values = palette_cat[1:3]) +
  labs(title = "MPG vs Weight, Colored by Cylinders, Sized by Horsepower",
       x = "Weight (1000 lbs)", y = "Miles per Gallon",
       color = "Cylinders", size = "Horsepower") +
  theme_minimal()
```

## Faceted Plots

```{r faceted-plot}
# Faceting splits the plot into small multiples by a categorical variable,
# making it easy to compare relationships across subgroups
ggplot(mtcars_df, aes(x = wt, y = mpg, color = cyl_f)) +
  geom_point(size = 2.5, alpha = 0.8) +
  geom_smooth(method = "lm", se = FALSE, color = main_color) +
  facet_wrap(~ am_f) +
  scale_color_manual(values = palette_cat[1:3]) +
  labs(title = "MPG vs Weight, Faceted by Transmission Type",
       x = "Weight (1000 lbs)", y = "Miles per Gallon", color = "Cylinders") +
  theme_minimal()
```

## 3D Scatter Plots (Static)

```{r scatter3d-static}
# Static 3D scatter plot using the scatterplot3d package
colors_3d <- palette_cat[as.numeric(mtcars_df$cyl_f)]
s3d <- scatterplot3d::scatterplot3d(
  x = mtcars_df$wt, y = mtcars_df$qsec, z = mtcars_df$mpg,
  color = colors_3d, pch = 19, angle = 55,
  main = "3D Scatter Plot: Weight, Quarter-Mile Time, and MPG",
  xlab = "Weight (1000 lbs)", ylab = "Quarter-Mile Time (sec)", zlab = "MPG"
)
legend(s3d$xyz.convert(4.5, 22, 32), legend = levels(mtcars_df$cyl_f),
       col = palette_cat[1:3], pch = 19, bty = "n", xpd = TRUE)
```

## Parallel Coordinates Plot

```{r parallel-coordinates}
# Parallel coordinates plot: each line is a car, each axis a standardized variable
# Useful for spotting clusters and outliers across many dimensions at once
MASS::parcoord(
  mtcars_df[, c("mpg", "disp", "hp", "drat", "wt", "qsec")],
  col = palette_cat[as.numeric(mtcars_df$cyl_f)],
  var.label = TRUE,
  main = "Parallel Coordinates Plot of Numeric Variables (colored by cylinders)"
)
legend("topright", legend = levels(mtcars_df$cyl_f), col = palette_cat[1:3],
       lty = 1, bty = "n", cex = 0.8)
```

------------------------------------------------------------------------

# Advanced Visualizations

## Principal Component Analysis (PCA) with Biplots

```{r pca-computation}
# Run PCA on standardized numeric variables (scale = TRUE standardizes units)
pca_vars <- mtcars_df %>% select(mpg, disp, hp, drat, wt, qsec)
pca_result <- prcomp(pca_vars, center = TRUE, scale. = TRUE)

summary(pca_result)
```

### Base R Biplot

```{r pca-biplot-base-r}
# Base R biplot: shows observations (points) and variable loadings (arrows)
biplot(pca_result,
       main = "PCA Biplot (Base R): mtcars Numeric Variables",
       col = c("#2980B9", "#E67E22"),
       cex = c(0.7, 0.9))
```

### ggplot2 Biplot

```{r pca-biplot-ggplot2}
# Manually build a ggplot2 biplot from PCA scores and loadings
pca_scores <- as.data.frame(pca_result$x) %>%
  mutate(cyl_f = mtcars_df$cyl_f, car = rownames(mtcars_df))

loadings <- as.data.frame(pca_result$rotation) %>%
  mutate(variable = rownames(pca_result$rotation))

# Scale factor to make loading arrows visible alongside score points
scale_factor <- 5

ggplot() +
  geom_point(data = pca_scores, aes(x = PC1, y = PC2, color = cyl_f), size = 3, alpha = 0.8) +
  geom_segment(data = loadings,
               aes(x = 0, y = 0, xend = PC1 * scale_factor, yend = PC2 * scale_factor),
               arrow = arrow(length = unit(0.25, "cm")), color = main_color) +
  geom_text(data = loadings,
            aes(x = PC1 * scale_factor * 1.15, y = PC2 * scale_factor * 1.15, label = variable),
            color = main_color, fontface = "bold", size = 3.5) +
  scale_color_manual(values = palette_cat[1:3]) +
  labs(title = "PCA Biplot (ggplot2): mtcars Numeric Variables",
       x = paste0("PC1 (", round(summary(pca_result)$importance[2,1] * 100, 1), "% variance)"),
       y = paste0("PC2 (", round(summary(pca_result)$importance[2,2] * 100, 1), "% variance)"),
       color = "Cylinders") +
  theme_minimal() +
  coord_equal()
```

## Clustered Heatmap

```{r clustered-heatmap, fig.width=8, fig.height=7}
# Clustered heatmap: rows (cars) and columns (variables) reordered by
# hierarchical clustering to reveal groupings/patterns in the data
scaled_mtcars <- scale(mtcars_df %>% select(mpg, disp, hp, drat, wt, qsec))

pheatmap(
  scaled_mtcars,
  scale = "none",                 # already scaled above
  clustering_method = "complete",
  color = colorRampPalette(c("#2980B9", "white", "#C0392B"))(100),
  main = "Clustered Heatmap of Standardized mtcars Variables",
  fontsize_row = 7,
  fontsize_col = 10
)
```

------------------------------------------------------------------------

# Interactive Visualizations

## Interactive 3D Scatter Plot (plotly)

```{r plotly-3d-scatter}
# Interactive 3D scatter plot: rotate/zoom with the mouse, hover for details
plot_ly(
  data = mtcars_df,
  x = ~wt, y = ~qsec, z = ~mpg,
  color = ~cyl_f, colors = palette_cat[1:3],
  text = ~paste("Car:", rownames(mtcars_df),
                "<br>HP:", hp,
                "<br>Transmission:", am_f),
  hoverinfo = "text",
  type = "scatter3d", mode = "markers",
  marker = list(size = 5, opacity = 0.85)
) %>%
  layout(
    title = "Interactive 3D Scatter: Weight, Quarter-Mile Time, MPG",
    scene = list(
      xaxis = list(title = "Weight (1000 lbs)"),
      yaxis = list(title = "Quarter-Mile Time (sec)"),
      zaxis = list(title = "Miles per Gallon")
    )
  )
```

## Interactive 2D Scatter Plot with Hover Information

```{r plotly-2d-scatter}
# Interactive 2D scatter plot built by converting a ggplot2 object with ggplotly,
# preserving rich hover tooltips with multiple variables
p_interactive <- ggplot(mtcars_df,
                         aes(x = wt, y = mpg, color = cyl_f,
                             text = paste("Car:", rownames(mtcars_df),
                                          "<br>Weight:", wt,
                                          "<br>MPG:", mpg,
                                          "<br>HP:", hp))) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(values = palette_cat[1:3]) +
  labs(title = "Interactive Scatter: MPG vs Weight",
       x = "Weight (1000 lbs)", y = "Miles per Gallon", color = "Cylinders") +
  theme_minimal()

ggplotly(p_interactive, tooltip = "text")
```

------------------------------------------------------------------------

# Summary

## Visualization Types by Analysis Category

```{r summary-table-viz-types}
viz_summary <- data.frame(
  `Analysis Type` = c(
    "Univariate (Numerical)", "Univariate (Numerical)", "Univariate (Numerical)",
    "Univariate (Categorical)", "Univariate (Categorical)",
    "Bivariate", "Bivariate", "Bivariate",
    "Multivariate", "Multivariate", "Multivariate", "Multivariate",
    "Advanced", "Advanced",
    "Interactive", "Interactive"
  ),
  `Visualization` = c(
    "Histogram + Density", "Boxplot + Jitter", "Q-Q Plot",
    "Bar Plot", "Pie Chart",
    "Scatter Plot + Regression", "Grouped Boxplot/Violin", "Mosaic Plot",
    "Scatterplot Matrix", "Correlation Heatmap", "3D Scatter", "Parallel Coordinates",
    "PCA Biplot", "Clustered Heatmap",
    "Interactive 3D Scatter (plotly)", "Interactive 2D Scatter (plotly)"
  ),
  `Best Used For` = c(
    "Shape/spread of a single variable", "Outliers and spread", "Assessing normality",
    "Comparing category counts", "Simple part-to-whole (use sparingly)",
    "Relationship between 2 numeric vars", "Comparing distributions across groups", "Relationship between 2 categorical vars",
    "Overview of all pairwise relationships", "Spotting strong/weak correlations fast", "Relationships among 3 numeric vars", "Patterns across many numeric vars",
    "Dimensionality reduction + variable structure", "Row/column clustering patterns",
    "Exploring 3D relationships interactively", "Exploring 2D relationships with detail on hover"
  ),
  check.names = FALSE
)

kable(viz_summary, caption = "Summary of Visualization Types by Analysis Category") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

## Best Practices Checklist

-   **Choose the right chart for the data type**: numeric distributions
    call for histograms/density/boxplots; categorical data calls for bar
    plots rather than pie charts.
-   **Always label axes and include units** so plots are interpretable
    without external context.
-   **Use color purposefully**, not decoratively — encode a meaningful
    variable (e.g., group membership) and keep palettes
    colorblind-friendly.
-   **Avoid 3D pie charts and unnecessary chart junk** — they distort
    perception of proportions and add no informational value.
-   **Show the underlying data when possible** (e.g., jittered points on
    boxplots) rather than summary statistics alone.
-   **Annotate statistical results directly on the plot** (correlation
    coefficients, p-values) to aid interpretation.
-   **Check assumptions before modeling** — use Q-Q plots and normality
    tests before assuming normality in downstream analyses.
-   **Standardize/scale variables** before multivariate techniques like
    PCA or clustering, since these methods are sensitive to differing
    variable scales.
-   **Use faceting instead of overplotting** when comparing more than a
    couple of subgroups.
-   **Reserve interactivity for exploration**, and provide static,
    exportable versions of key figures for reports and publications.

## Key R Packages Used

```{r summary-table-packages}
package_summary <- data.frame(
  Package = c("ggplot2", "dplyr / tidyr", "moments", "GGally", "corrplot",
              "vcd", "scatterplot3d", "MASS", "pheatmap", "plotly", "knitr / kableExtra"),
  Purpose = c(
    "Grammar-of-graphics static plotting (histograms, boxplots, scatter, facets, etc.)",
    "Data wrangling, grouping, and reshaping prior to visualization",
    "Skewness and kurtosis calculations for distribution shape",
    "Scatterplot matrices (ggpairs) for multivariate exploration",
    "Correlation matrix visualizations (heatmaps with values)",
    "Mosaic plots for categorical x categorical relationships",
    "Static 3D scatter plots in base graphics",
    "Parallel coordinates plots (parcoord) and other multivariate tools",
    "Clustered heatmaps with hierarchical clustering dendrograms",
    "Interactive 2D/3D plots with zoom, pan, and hover tooltips",
    "Rendering clean, styled summary tables in HTML output"
  )
)

kable(package_summary, caption = "Key R Packages Used in This Document") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

------------------------------------------------------------------------

# Session Info

```{r session-info}
# Full session info for reproducibility (R version, OS, and all loaded packages)
sessionInfo()
```
