---
title: "Assignment II"
author: "Gershon Tetteh Amanor"
date: "`r Sys.Date()`"
output:
  #word_document: default
  html_document: default
---

| NAMES                 | IDS      |
|-----------------------|----------|
| GERSHON TETTEH AMANOR | 10967696 |
| JOHN                  |          |
| SYLVESTER             |          |
| SYLVESTER             |          |
|                       |          |

: GROUP MEMBERS

------------------------------------------------------------------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
# loading the MASS package since the anorexia data is in the MASS package
library(MASS)

# setting the margins
par(mar = c(4,4,1,1))
boxplot(Prewt~Treat,
        data = anorexia,
        xlab = 'Treatment Type',
        ylab = 'Pre weight (lbs)',
        col = c('red','white','blue'))
```

## R Markdown

The anorexia data is a Weight change data for 72 young female anorexia patients. Here we want to compare the distribution of the pre-weight of the patients on the three different treatment types, **"Cont"** (control), **"CBT"** (Cognitive Behavioural treatment) and **"FT"** (family treatment).

plan

1.  Looking at the shapes of these plots, *CBT* and *FT* are slightly normally distributed since the lengths of the upper and lower whiskers are slightly the same for both of them but that of *Cont* is slightly positively skewed.

2.  The *CBT* group has two extreme weights (outliers) of about 70lbs and 95lbs however, the *Cont* and *FT* groups have no extreme weights.

3.  on average, the *CBT* and *FT* groups have equal weights of about 83lbs however, the *Cont* group on average weighs about 80lbs.

4.  spread

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
boxplot(count~spray,
        data=InsectSprays,
        col = rainbow(6),
        ylab = 'count of insects',
        xlab = 'spray')
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
